The relationship between age group(15-24, 24-54,55 and over) with the unemployment rate.
The relationship between sex group(Females,males) with the unemployment rate. 



```{r}
#rdata = read.csv('emp_basic_data.csv')
library(readr)
rdata <- read_csv("emp_basic_data.csv")
#View(rdata)
```
```{r}
tail(rdata,4)
```
First, display the bar chart for Sex group females
```{r}
rdata$REF_DATE <- as.character(rdata$REF_DATE)

data.2000to2020 = filter(rdata,REF_DATE >= '2000-01', REF_DATE<='2020-08')
colnames(data.2000to2020)[4] = "Labour.force.characteristics"
colnames(data.2000to2020)[6] = "Age.group"
data.2000to2020 
```
Filter out the value for labour force characteristics as we only need unemployment rate and females.

plot only Canada female unemployment rate to see the normalbility, from the table below, the nsize for this certain condistion is 248.

```{r}
data.2000to2020.CA = select(filter(data.2000to2020,GEO=='Canada'),REF_DATE,GEO,Labour.force.characteristics,Sex,Age.group,VALUE)
data.2000to2020.CA = select(filter(data.2000to2020.CA,Age.group=='15 years and over'),REF_DATE,GEO,Sex,Labour.force.characteristics,VALUE,Age.group)
data.2000to2020.CA = select(filter(data.2000to2020.CA,Sex=='Females'|Sex=='Males'),REF_DATE,GEO,Sex,Labour.force.characteristics,VALUE,Age.group)
data.2000to2020.CA = select(filter(data.2000to2020.CA,Labour.force.characteristics=='Unemployment rate'),REF_DATE,GEO,Sex,Labour.force.characteristics,VALUE,Age.group)
data.2000to2020.CA
```
For all provinces and Canada:
```{r}
data.2000to2020.all = select(filter(data.2000to2020,Age.group=='15 years and over'),REF_DATE,GEO,Sex,Labour.force.characteristics,VALUE,Age.group)
data.2000to2020.all
```
```{r}
ggplot(data=filter(data.2000to2020.CA,Sex=='Females'),aes(x=VALUE))+ geom_histogram(col="red", fill="blue", binwidth=0.5)+xlab('Unemployment rate(%)')+ylab('Count Frequency for unemployment rate')+ggtitle('Histogram of Unemployment rate for Female')
```
The graph is right-skewed. 
```{r}
ggplot(data=filter(data.2000to2020.CA,Sex=='Males'),aes(x=VALUE))+ geom_histogram(col="red", fill="blue", binwidth=0.5)+xlab('Unemployment rate(%)')+ylab('Count Frequency for unemployment rate')+ggtitle('Histogram of Unemployment rate for Males')
```
The graph is right-skewed. 

For more clear stas analysis, the boxplot and violin plot is below:

```{r}
ggplot(data.2000to2020.CA,aes(x=Sex,y=VALUE))+ geom_violin(col="blue", fill="red", na.rm=TRUE) +xlab('Sex')+ geom_boxplot(binwidth=0.5,ns.rm=TRUE)+ylab('Unemployment rate(%)')+ggtitle('Box plot of Unemployment rate for Female and Male')+ coord_flip()
```
```{r}
data.2000to2020.all= select(filter(data.2000to2020,Sex=='Females'|Sex=='Males' ),REF_DATE,GEO,Sex,Labour.force.characteristics,VALUE,Age.group)
data.2000to2020.all= select(filter(data.2000to2020.all,Labour.force.characteristics=='Unemployment rate' ),REF_DATE,GEO,Sex,Labour.force.characteristics,VALUE,Age.group)
tail(data.2000to2020.all,4)
```
```{r}
ggplot(data=filter(data.2000to2020.all,Sex=='Females',GEO!='Canada'),mapping = aes(x=VALUE))+geom_histogram(col='blue', binwidth=1, na.rm=TRUE) + xlab("Unemployment Rate(%)") + facet_wrap(~GEO)+ggtitle('Unemployment rate for Different Canada Provinces(Females)')
```
```{r}
ggplot(data=filter(data.2000to2020.all,Sex=='Males',GEO!='Canada'),mapping = aes(x=VALUE))+geom_histogram(col='blue', binwidth=1, na.rm=TRUE) + xlab("Unemployment Rate(%)") + facet_wrap(~GEO)+ggtitle('Unemployment rate for Different Canada Provinces(Males)')
```
Now, for data from Canada, Can we conclude the mean of unemployment for women is larger than the mean of unemployment for man?

First, use the permutation method to get the diffrence between males and females, use N=10000 for the bootstrap
```{r}
data.2000to2020.CA
```
```{r}
favstats(~VALUE|Sex,data=data.2000to2020.CA)
```

$$
{\rm H}_{0}: \mu_{M} \leq \mu_{F} \hspace{0.25in} {\rm } \hspace{0.25in} {\rm H}_{A}: \mu_{M} > \mu_{F} 
$$
The Null Hypothese is the mean of Females is smaller or equal to mean of Males for unemployment rate. 
The Alternative Hypothese is the mean of Males is larger than mean of Females for unemployment rate.
```{r}
fmdiff = mean(~VALUE,data=filter(data.2000to2020.CA,Sex=='Females')) - mean(~VALUE,data=filter(data.2000to2020.CA,Sex=='Males'))
fmdiff
n.all = dim(data.2000to2020.CA)[1]
n.all
data.2000to2020.CAf = filter(data.2000to2020.CA,Sex=='Females')
n.female = dim(data.2000to2020.CAf)[1]
n.female
n.males = 496-248
```
```{r}
N=2000
fmoutcome = numeric(N)

for( i in 1:N){
  index.1 = sample(496,248,replace=FALSE)
  fmoutcome[i] = mean(data.2000to2020.CA$VALUE[index.1])-mean(data.2000to2020.CA$VALUE[-index.1])
}

```
```{r}
hist(fmoutcome,xlab="Differnce of Unemployment of Females and Males",ylab="Frequency",main="Outcome of 10000 Permulation Tests",col='blue')
abline(v=fmdiff,col="red")
```
```{r}
(sum(fmoutcome<=(fmdiff)))/N
```
The null hyposes will be rejected. So that the mean of unemployment rate for female is less than the mean of unemployment rate for male.

How about Parametric Testing?
First, cheeck the normality conditon.

```{r}
ggplot(data=filter(data.2000to2020.CA,Sex=='Females'),aes(sample=VALUE))+ stat_qq(size=2, col='blue') + stat_qqline(col='red') + ggtitle("Normal Probability Check for Females")
```
```{r}
ggplot(data=filter(data.2000to2020.CA,Sex=='Males'),aes(sample=VALUE))+ stat_qq(size=2, col='blue') + stat_qqline(col='red') + ggtitle("Normal Probability Check for Males")
```

```{r}
pnorm(0.8)
```




Data wrangling
```{r}
oilprice = read.csv("oilprice.csv")
oilprice$YYMM <- as.character(oilprice$YYMM)
oildata.2000to2020 = filter(oilprice,YYMM >= '2000-01'& YYMM<='2020-08')
names(oildata.2000to2020)[1] = "REF_DATE"
oildata.2000to2020
oildataLR = filter(rdata,REF_DATE >= '2000-01'& REF_DATE<='2020-08')
oildataLR=data.frame(oildataLR)
dataemp = select(filter(oildataLR,GEO=='Canada',Sex =='Both sexes',Labour.force.characteristics
=='Unemployment rate',Age.group=='15 years and over'),REF_DATE,VALUE)
totaldataLR = merge(oildata.2000to2020,dataemp,by="REF_DATE")
totaldataLR
ggplot(data=totaldataLR, aes(x = Price, y = VALUE)) + geom_point(col="blue", size=2, position="jitter") + ylab("Unemployment rate (%)") + xlab("Oilprice") + ggtitle("Scatterplot of Unemployment rate (%) to Oil Price")+ stat_smooth(method="lm", col='red')

```

```{r}
dataempAB = select(filter(oildataLR,GEO=='Alberta',Sex =='Both sexes',Labour.force.characteristics
=='Unemployment rate',Age.group=='15 years and over'),REF_DATE,VALUE)
totaldataLRAB = merge(oildata.2000to2020,dataempAB,by="REF_DATE")
totaldataLRAB
```


```{r}
```

$$
\begin{align}
{\rm H}_{0}:& \text{the Y-variable CANNOT be expressed as a linear function of X-variable} \\
{\rm H}_{A}:& \text{the Y-variable CAN be expressed as a linear function of X-variable} \\
\end{align}
$$

```{r}
LRAB = data.frame(Price =totaldataLRAB$Price,UnemploymentRate = totaldataLRAB$VALUE, REF_DATE = totaldataLRAB$REF_DATE)
#LRAB = tally(~Price + UnemploymentRate, LRAB,margin = FALSE)
LRAB
```

```{r}
ggplot(data=LRAB, aes(x = Price, y = UnemploymentRate)) + geom_point(col="blue", size=2, position="jitter") + ylab("Unemployment rate (%)") + xlab("oil price") + ggtitle("Scatterplot of Unemployment rate (%) to oil price")+ stat_smooth(method="lm", col='red')
cor(~Price,~UnemploymentRate, data = LRAB)
predunemp = lm(UnemploymentRate~Price, LRAB)
options(scipen=999)
predunemp$coef
```
```{r}
predictsUempei = predunemp$fitted.values 
eisUnemp = predunemp$residuals    
diagnosticdf = data.frame(predictsUempei, eisUnemp)
ggplot(diagnosticdf, aes(sample = eisUnemp)) +  stat_qq(col='blue') + stat_qqline(col='red') + ggtitle("Normal Probability Plot of Residuals")
ggplot(diagnosticdf, aes(x = predictsUempei, y = eisUnemp)) +  geom_point(size=2, col='blue', position="jitter") + xlab("Predicted balance") + ylab("Residuals") + ggtitle("Plot of Fits to Residuals") + geom_hline(yintercept=0, color="red", linetype="dashed")
summary(aov(predunemp))
rsquared(predunemp)
```

```{r}
Nbootstraps6 = 1000 
cor.boot6 = numeric(Nbootstraps6)
a.boot6 = numeric(Nbootstraps6) 
b.boot6 = numeric(Nbootstraps6) 
ymean.boot6 = numeric(Nbootstraps6)
rsquare.boot6=numeric(Nbootstraps6)
nsize6 =  dim(LRAB)[1]
xvalue6 = 60000
for(i in 1:Nbootstraps6) 
{ index = sample(nsize6, replace =TRUE)
LRAB.boot = LRAB[index,]
cor.boot6[i]=cor(~Price,~UnemploymentRate,data =LRAB.boot)
LRAB.lm = lm(UnemploymentRate~Price, data=LRAB.boot)  
    a.boot6[i] = coef(LRAB.lm)[1] 
    b.boot6[i] = coef(LRAB.lm)[2] 
    ymean.boot6[i] = a.boot6[i] + (b.boot6[i]*xvalue6)
    rsquare.boot6[i]=rsquared(LRAB.lm)
} 

bootstraplrab = data.frame(cor.boot6,a.boot6,b.boot6,rsquare.boot6) 
head(bootstraplrab,3)
favstats(~a.boot6, data=bootstraplrab)
favstats(~b.boot6, data=bootstraplrab)
favstats(~rsquare.boot6, data=bootstraplrab)

ggplot(bootstraplrab, aes(x = cor.boot6)) + geom_histogram(col="red", fill="blue", binwidth=0.05) + xlab("Values of the Bootstrap Statistic: Correlation Coefficient") + ylab("Count") + ggtitle("Distribution of Bootstrap Statistics: r")
qdata(~cor.boot6,c(0.025,0.975),data = bootstraplrab)
```


what about Gender ratio?

```{r}

#ggplot(data.2000to2020.CA, aes(x = Sex, y = VALUE)) + geom_point(col="blue", size=2, position="jitter") + ylab("Unemployment rate") + xlab("Sex") + ggtitle("Scatterplot of balance to income")+ stat_smooth(method="lm", col='red')

femaledataca =select(filter(oildataLR,GEO=='Canada',Sex =='Females',Labour.force.characteristics
=='Population',Age.group=='15 years and over'),REF_DATE,VALUE)
names(femaledataca)[2] = "female.population"
maledataca =select(filter(oildataLR,GEO=='Canada',Sex =='Males',Labour.force.characteristics
=='Population',Age.group=='15 years and over'),REF_DATE,VALUE)
names(maledataca)[2] = "male.population"
Ratioca = femaledataca$female.population/maledataca$male.population
femaledataca$Ratio=Ratio
Ratiodataca = merge(femaledataca,totaldataLR,by="REF_DATE")
Ratiodataca
ggplot(Ratiodataca, aes(x = Ratio, y = VALUE)) + geom_point(col="blue", size=2, position="jitter") + ylab("Unemployment rate (%)") + xlab("Gender Raio(F/M)") + ggtitle("Scatterplot of Unemployment rate (%) to Gender Raio(F/M)")+ stat_smooth(method="lm", col='red')
```

Alberta Ratio to unemployment rate
```{r}
#ggplot(data.2000to2020.CA, aes(x = Sex, y = VALUE)) + geom_point(col="blue", size=2, position="jitter") + ylab("Unemployment rate") + xlab("Sex") + ggtitle("Scatterplot of balance to income")+ stat_smooth(method="lm", col='red')

femaledata =select(filter(oildataLR,GEO=='Alberta',Sex =='Females',Labour.force.characteristics
=='Population',Age.group=='15 years and over'),REF_DATE,VALUE)
names(femaledata)[2] = "female.population"
maledata =select(filter(oildataLR,GEO=='Alberta',Sex =='Males',Labour.force.characteristics
=='Population',Age.group=='15 years and over'),REF_DATE,VALUE)
names(maledata)[2] = "male.population"
Ratio = femaledata$female.population/maledata$male.population
femaledata$Ratio=Ratio
Ratiodata = merge(femaledata,LRAB,by="REF_DATE")
Ratiodata
ggplot(Ratiodata, aes(x = Ratio, y = UnemploymentRate)) + geom_point(col="blue", size=2, position="jitter") + ylab("Unemployment rate (%)") + xlab("Gender Raio(F/M)") + ggtitle("Scatterplot of Unemployment rate (%) to Gender Raio(F/M)")+ stat_smooth(method="lm", col='red')
```

What about age?
```{r}
age15dataca =select(filter(oildataLR,GEO=='Canada',Sex =='Both sexes',Labour.force.characteristics
=='Population',Age.group=='15 to 24 years'),REF_DATE,VALUE)
names(age15dataca)[2] = "15-19.population"
age25dataca =select(filter(oildataLR,GEO=='Canada',Sex =='Both sexes',Labour.force.characteristics
=='Population',Age.group=='25 to 54 years'),REF_DATE,VALUE)
names(age25dataca)[2] = "25-54.population"
age55dataca =select(filter(oildataLR,GEO=='Canada',Sex =='Both sexes',Labour.force.characteristics
=='Population',Age.group=='55 to 64 years'),REF_DATE,VALUE)
names(age55dataca)[2] = "55-64.population"
totalpopulation = select(filter(oildataLR,GEO=='Canada',Sex =='Both sexes',Labour.force.characteristics
=='Population',Age.group=='15 years and over'),REF_DATE,VALUE)
totalpopulation
p15 = age15dataca$`15-19.population`/totalpopulation$VALUE
p25 = age25dataca$`25-54.population`/totalpopulation$VALUE
p55 = age55dataca$`55-64.population`/totalpopulation$VALUE

age15dataca$p15=p15
age25dataca$p25=p25
age55dataca$p55=p55
p15dataca = merge(age15dataca,totaldataLR,by="REF_DATE")
p25dataca = merge(age25dataca,totaldataLR,by="REF_DATE")
p55dataca = merge(age55dataca,totaldataLR,by="REF_DATE")
ggplot(p15dataca, aes(x = p15, y = VALUE)) + geom_point(col="blue", size=2, position="jitter") + ylab("Unemployment rate (%)") + xlab("Gender Raio(F/M)") + ggtitle("Scatterplot of Unemployment rate (%) to Gender Raio(F/M)")+ stat_smooth(method="lm", col='red')
ggplot(p25dataca, aes(x = p25, y = VALUE)) + geom_point(col="blue", size=2, position="jitter") + ylab("Unemployment rate (%)") + xlab("Gender Raio(F/M)") + ggtitle("Scatterplot of Unemployment rate (%) to Gender Raio(F/M)")+ stat_smooth(method="lm", col='red')
ggplot(p55dataca, aes(x = p55, y = VALUE)) + geom_point(col="blue", size=2, position="jitter") + ylab("Unemployment rate (%)") + xlab("Gender Raio(F/M)") + ggtitle("Scatterplot of Unemployment rate (%) to Gender Raio(F/M)")+ stat_smooth(method="lm", col='red')
```
```{r}
MRLDATA = cbind(p15dataca$p15,p25dataca$p25,p55dataca$p55,Ratiodata$Ratio,totaldataLR)
MRLDATA
model <- lm(MRLDATA$VALUE ~ p15dataca$p15 + Ratiodata$Ratio+MRLDATA$Price)
summary(model)
rsquared(model)
```

